A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation

### 2016-06-20 talk_Roscoff.tex
backslides

\begin{frame}{Gel formation}
	\begin{columns}
		\column{0.5\textwidth}
		\begin{tikzpicture}[
			every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=1 by 2,
					horizontal sep=4em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
				},
			cycle list name=samples,
			width=\textwidth,
			%height=0.5\linewidth,
			axis on top,
		]
		\nextgroupplot[
			ylabel={$\Pi v$},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd};
			\addplot[black, no marks] file {res_gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
			\node at (axis cs:0.4,1500) {Gel};

		\nextgroupplot[
			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd};
			\addplot[black, no marks] file {gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
			\draw[->] (x.base) --(rel axis cs:0.98,0.01);
		\end{groupplot}
		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
		\end{tikzpicture}
		\column{0.5\textwidth}
		Small polymer/colloid size ratio
		\begin{itemize}
		\item metastable gas-liquid
		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
		\item continuous minority phase
		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
		\end{itemize}
		\begin{tikzpicture}
	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.1, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.5, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|};
	\fill[ilmcolor] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle ($(t1)!0.9!45:(t2)$) coordinate(cnw);
	\clip (cse) rectangle (cnw);
	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
		+(60:0.06\textwidth) circle (0.045\textwidth) 
		+(180:0.06\textwidth) circle (0.045\textwidth) 
		+(-60:0.06\textwidth) circle (0.045\textwidth) 
		+(0:0.12\textwidth) circle (0.045\textwidth)
		+(120:0.12\textwidth) circle (0.045\textwidth)
		+(-120:0.12\textwidth) circle (0.045\textwidth)
		+(45:0.16\textwidth) circle (0.045\textwidth)
		+(-45:0.16\textwidth) circle (0.045\textwidth);
	\end{tikzpicture}
	\end{columns}
\end{frame}

@@

>>>>

\begin{frame}{Gel formation}
	\begin{columns}
		\column{0.5\textwidth}
		\vspace{\baselineskip}
		\tikzsetnextfilename{phdiagth}%
		\begin{tikzpicture}[
			every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=1 by 2,
					horizontal sep=4em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
				},
			cycle list name=samples,
			width=\textwidth,
			%height=0.5\linewidth,
			axis on top,
		]
		\nextgroupplot[
			ylabel={$\Pi v$},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd};
			\addplot[black, no marks] file {res_gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd} node[above left, inner ysep=0, inner xsep=1pt] (x) at (rel axis cs:1,0) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
			\node at (axis cs:0.4,1500) {Gel};

		\nextgroupplot[
			xlabel={$\phi$}, ylabel={$c_p$ (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd};
			\addplot[black, no marks] file {gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} node[above right] at (rel axis cs:0,0) {Fluid} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd} node[above left] (x) at (rel axis cs:1,0.5) {Crystal}  -| (rel axis cs:1,0) \closedcycle;
			\draw[->] (x.base) --(rel axis cs:0.98,0.01);
		\end{groupplot}
		%\draw(g c1r1.outer south west) -- +(\textwidth,0);
		\end{tikzpicture}
		\column{0.5\textwidth}
		Small polymer/colloid size ratio
		\begin{itemize}
		\item metastable gas-liquid
		\item spinodal + glass $\Rightarrow$ gel\\ \hfill\textit{\footnotesize Lu et al. Nature~2008}
		\item continuous minority phase
		\item viscoelastic phase sep.\\ \hfill\textit{\footnotesize Tanaka J.Phys.Cond.Mat.~2000}
		\item due to dynamic assymetry
		\end{itemize}
		\tikzsetnextfilename{vps}%
		\begin{tikzpicture}
	\draw[->] (0,0) -- +(\textwidth,0) node[pos=0.05, label=-90:\textcolor{ilmorange}{$\tau_1$}](t1) {|} node[pos=0.54, label=-90:\textcolor{ilmcolor}{$\tau_2$}] (t2){|} coordinate[pos=1] (tf) node[anchor=north east,font=\footnotesize]{time};
	\let\side\relax
	\newlength\side
	\setlength{\side}{0.275\textwidth}
	\fill[ilmcolor!80!ilmorange] ($(t1)!0.1!45:(t2)$) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange!80!ilmcolor] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.22\side) 
		+(180:0.3\side) circle (0.22\side) 
		+(-60:0.3\side) circle (0.22\side) 
		+(0:0.6\side) circle (0.22\side)
		+(120:0.6\side) circle (0.22\side)
		+(-120:0.6\side) circle (0.22\side)
		+(45:0.8\side) circle (0.22\side)
		+(-45:0.8\side) circle (0.22\side);
	\end{scope}
	\fill[ilmcolor] (cse-|t2) ++(-0.5\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmorange] ++($(cse)!0.5!(cnw)$) 
		+(60:0.3\side) circle (0.26\side) 
		+(180:0.3\side) circle (0.26\side) 
		+(-60:0.3\side) circle (0.25\side) 
		+(0:0.6\side) circle (0.25\side)
		+(120:0.6\side) circle (0.25\side)
		+(-120:0.6\side) circle (0.25\side)
		+(45:0.8\side) circle (0.25\side)
		+(-45:0.8\side) circle (0.25\side)
		%enhance broken bond
		+(120:0.15\side) circle (0.07\side);
	\end{scope}
	\fill[ilmorange] (cse-|tf) ++(-\side,0) coordinate(cse) rectangle +(\side,\side) coordinate(cnw);
	\begin{scope}
	\clip (cse) rectangle (cnw);
	\fill[ilmcolor] ++($(cse)!0.5!(cnw)$) circle (0.07\side)
		+(0:0.3\side) circle (0.07\side) 
		+(120:0.3\side) circle (0.07\side) 
		+(-120:0.3\side) circle (0.07\side) 
		+(-60:0.6\side) circle (0.07\side)
		+(60:0.6\side) circle (0.07\side)
		++(0:0.3\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side)
		+(120:0.6\side) circle (0.07\side)
		+(-120:0.6\side) circle (0.07\side)
		++(180:0.9\side)
		+(-60:0.3\side) circle (0.07\side)
		+(60:0.3\side) circle (0.07\side);
	\end{scope}
	\end{tikzpicture}

	\hfill$\textcolor{ilmcolor}{\tau_2} \rightarrow +\infty$ {\usebeamercolor[fg]{itemize item}\usebeamertemplate{itemize item}} arrest 
	\end{columns}
\end{frame}

@@

\begin{frame}{Reproducibility in capillary}
	\hfill $\phi\approx 0.33$, $c_P=\SI{0.40}{\milli\gram/\gram}$

	\includegraphics[width=\textwidth]{nature/capillary}
\end{frame}

\begin{frame}{Transition probabilities}
	\vspace{\baselineskip}
	\includegraphics[width=\textwidth]{nature/sfig1.pdf}
\end{frame}

\begin{frame}{Evolution local volume fraction}
	\vspace{\baselineskip}
	12-neighboured particles
	\begin{center}
	\includegraphics{nature/local_phi12.pdf}
	\end{center}
\end{frame}

\begin{frame}{phase definition}
	\begin{description}
	\item[crystal]
		\begin{itemize} 
		\item particle $i$ has $N_b(i)=12$ neighbours
		\item $\displaystyle q_{6 m}(i)=\frac{1}{N_b(i)}\sum_{j=1}^{N_b(i)}
	Y_{6 m}(\mathbf{r_{ij}}),\quad -6 <m<6$
		\item crystalline bond $(i,j)$: \[(\mathbf{q}_{6}(i)/|\mathbf{q}_{6}(i)|)\cdot(\mathbf{q}_{6}(j)/|\mathbf{q}_{6}(j)|)>0.7\]
		\item crystalline particle: $7$ crystalline bonds
		\end{itemize}
	\item[surface] two neighbouring crystal particles
	\item[liquid] at least four neighbouring particles
	\item[gas]
	\end{description}
\end{frame}

\begin{frame}{Colloid and polymer sizes}
	Direct measurement \alert{too imprecise}
	\begin{itemize}
	\item diffusion constant $\Rightarrow R= \SI{2.3 +- 0.05}{\micro\metre}$
	\item Flory scaling litterature measurements $\Rightarrow R_g=\SI{76. +- 50.}{\nano\metre}$
	\end{itemize}

	\begin{block}{Adjusting theoretical and experimental phase diagrams}
	\begin{enumerate}
	\item to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction 
	\item to have the theoretical spinodal line to correspond to the gel region boundary.
	\end{enumerate}
	\end{block}
	Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=\SI{80}{\nano\metre}$, $\sigma = \SI{2.25}{\micro\metre}$ and a crystal (without defects) at $0.723$.
\end{frame}

\begin{frame}{Carnahan-Starling is not enough}
	\tikzsetnextfilename{CSphasediag}%
	\begin{tikzpicture}[
		every axis/.style={
			xlabel absolute, every axis x label/.append style={anchor=base, yshift=-0.5em},
			ylabel absolute, every axis y label/.append style={anchor=base, yshift=-0.5em}
			}]
		\pgfplotscreateplotcyclelist{samples}{%
		blue,mark=square*,\\%
		cyan,mark=square\\%
		red, thick, mark=*\\%
		red!65!yellow, mark=otimes\\%
		red!30!yellow, mark=oplus\\%
		red!65!black, mark=o\\%
		}%
		\begin{groupplot}[
			group style={
					group name=g, group size=2 by 2,
					horizontal sep=0.5em,
					vertical sep=0.5em,
					x descriptions at=edge bottom,
					y descriptions at=edge left,
				},
			cycle list name=samples,
			width=0.5\textwidth,
			%height=0.5\linewidth,
			axis on top,
			xlabel={$\phi$},
		]
		\nextgroupplot[
			ylabel={attraction},
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
		\addplot file {res_sample0.phd};
		\addplot file {res_sample1.phd};
		\addplot file {res_sample2.phd};
		\addplot file {res_sample3.phd};
		\addplot file {res_sample4.phd};
		\addplot file {res_sample5.phd};
		\end{scope}
			\addplot[black, dashed,no marks] file {res_gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquid_sl.phd};
			\addplot[black, no marks] file {res_gasliquid_bl.phd};
			\addplot[black, no marks] file {res_gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;

	\nextgroupplot[
			xmin=0,xmax=0.74,
			ymin=0, ymax=2400, %ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
		\addplot file {res_sample0.phd};
		\addplot file {res_sample1.phd};
		\addplot file {res_sample2.phd};
		\addplot file {res_sample3.phd};
		\addplot file {res_sample4.phd};
		\addplot file {res_sample5.phd};
		\end{scope}
			\addplot[black, dashed,no marks] file {res_gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {res_gasliquidCS_sl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {res_gasliquidCS_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {res_fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;

		\nextgroupplot[
			ylabel={P concentr. (\si{\milli\gram/\gram})},
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
			\addplot file {sample0.phd};
			\addplot file {sample1.phd};
			\addplot file {sample2.phd};
			\addplot file {sample3.phd};
			\addplot file {sample4.phd};
			\addplot file {sample5.phd};
			\end{scope}
			\addplot[black, dashed,no marks] file {gasliquid_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquid_sl.phd};
			\addplot[black, no marks] file {gasliquid_bl.phd};
			\addplot[black, no marks] file {gasliquid_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
			\addplot[only marks, mark=triangle] file{fluid_samples.phd};

		\nextgroupplot[
			xmin=0,xmax=0.74,
			ymin=0, ymax=1.9, ytick={0,0.5,...,2},
			mark options={solid}
			]
			\begin{scope}[dotted, mark repeat=2,mark phase=2,<->]
			\addplot file {sample0.phd};
			\addplot file {sample1.phd};
			\addplot file {sample2.phd};
			\addplot file {sample3.phd};
			\addplot file {sample4.phd};
			\addplot file {sample5.phd};
			\end{scope}
			\addplot[black, dashed,no marks] file {gasliquidCS_sg.phd};% node[diamond, fill, inner sep=2pt] at (current plot begin){};
			\addplot[black, dashed,no marks] file {gasliquidCS_sl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bl.phd};
			\addplot[black, no marks,skip coords between index={10}{100}] file {gasliquidCS_bg.phd};
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_f.phd} -- (rel axis cs:0,0) \closedcycle;
			\addplot[fill=lightgray!50, no marks, draw=none] file {fluidcrystal_x.phd}  -| (rel axis cs:1,0) \closedcycle;
			\addplot[only marks, mark=triangle] file{fluid_samples.phd};
	\end{groupplot}
	\end{tikzpicture}
\end{frame}
